---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MVMRcML

<!-- badges: start -->
<!-- badges: end -->

The goal of MVMRcML is to conduct Multivariable Mendelian Randomization (MVMR) analysis.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ZhaotongL/MVMR-cML")
```

## Example

Here is an example which shows how to apply MVMR-cML to infer the causal relationship from cholesterol levels (triglycerides, LDL and HDL) to CAD.

First extract GWAS summary data with \code{TwoSampleMR} package:
```{r}
library(TwoSampleMR)
gwas_id = c(
      'ebi-a-GCST002216',
      'ebi-a-GCST002222',
      'ebi-a-GCST002223')
exposure_dat = mv_extract_exposures(gwas_id)
outcome_dat <- extract_outcome_data(exposure_dat$SNP,'ebi-a-GCST005195')
mvdat <- mv_harmonise_data(exposure_dat, outcome_dat)
```

MVMR-cML analysis:
```{r}
library(MVMRcML)
## rho_mat is pre-calculated by bivariate LDSC
rho_mat = matrix(c(1,0.1988,-0.3347,0,0.1988,1,-0.1114,0,-0.3347,-0.1114,1,0,0,0,0,1),ncol=4)
Sig_inv_l = invcov_mvmr(se_bx=mvdat$exposure_se,se_by=mvdat$outcome_se,rho_mat = rho_mat)

MVcML_res = MVmr_cML_DP(b_exp=mvdat$exposure_beta,
                        b_out=as.matrix(mvdat$outcome_beta),
                        se_bx=mvdat$exposure_se,
                        Sig_inv_l=Sig_inv_l,n = 188577,num_pert = 100,
                        K_vec = 0:20 # try a small range of K first to save time. Change this accordingly.
                        )
```

In above code, we first tried a small range of $K$ from 0 to 20 (which can be changed accordingly in practice), and MVMR-cML-BIC selected 8 invalid IVs.
```{r}
## Invalid IVs selected by MVMR-cML-BIC
MVcML_res$BIC_invalid

## Inference by MVMR-cML-BIC (NOT ACCOUNTING FOR MODEL SELECTION UNCERTAINTY!)
MVcML_BIC_SE = MVcML_SdTheta(b_exp=mvdat$exposure_beta,
              b_out=as.matrix(mvdat$outcome_beta),
              Sig_inv_l=Sig_inv_l,
              theta=MVcML_res$BIC_theta,
              zero_ind = setdiff(1:length(mvdat$outcome_beta),MVcML_res$BIC_invalid))

MVcMLBIC_pval = pnorm(-abs(MVcML_res$BIC_theta/MVcML_BIC_SE))*2
```

MVMR-cML-DP suggested significant direct effects of TG and LDL on CAD, and a null effect of HDL on CAD.
```{r}
## Inference by MVMR-cML-DP
MVcMLDP_pval = pnorm(-abs(MVcML_res$BIC_DP_theta/MVcML_res$BIC_DP_se))*2
MVcMLDP_pval
```
